Box 1 - TDT Curves

### 2023-09-21 ms/ms.qmd
More writing

```{r, newspp}
#| label: newspp
#| echo: false
#| message: false
#| include: false

@@

```{r, fig-tdt2}
```

@@

A number of significance challenges imped are ability to take the TDT and sub-lethal frameworks into a comparative context. First, ethical, conservation, and logistical challenges limit our ability to measure and estimate key parameters for many species, creating significant gaps in our knowledge across the tree of life. Second, while there has been a focus on species-level responses, population-level information is seldom incoperated into comparative frameworks.

Powerful multivariate hierachical models (i.e., multi-response models) are capable of dealing with missing data, and provide a means by which multiple physiological thermal tolerance measures can be simultaneously modelled. When used in conjunction with phylogenetic information across species and/or populations, such modelling approaches can provide informed estimates for species missing data and explicitly estimate the covariance between tolerance measures while accounting for their uncertainty. Such insight that can be used to inform policy and conservation-related decision making.

Using a well studied taxonomic group, the *Drosophila* clade, we demonstrate how such an analytical approach can inform on these significant questions.

>>>>

## Estimate TDT Curve Parameters for *Drosophilia*
```{r, newspp}
#| label: newspp
#| echo: false
#| message: false
#| include: false

@@

## Applying Multivariate Modelling Approaches

@@

A number of significance challenges imped are ability to take the TDT and sub-lethal frameworks into a comparative context. First, ethical, conservation, and logistical challenges limit our ability to measure and estimate key parameters for many species, creating significant gaps in our knowledge across the tree of life. Second, while there has been a focus on species-level responses, population-level information is seldom incoperated into comparative frameworks limiting our ability to decompose sources of variance.

Powerful multivariate hierachical models (i.e., multi-response models) are capable of dealing with missing data, and provide a means by which multiple physiological thermal tolerance measures can be modelled simultaneously [@Noble2021; @Pottier2023]. When used in conjunction with an understanding of the phylogenetic relationships among species and/or populations, such modelling approaches can provide informed estimates for species missing data and explicitly estimate the covariance between tolerance measures while accounting for their uncertainty. Such insight that can be used to inform policy and conservation-related decision making for species with little to no data.

#### Application of Multivariate Comparative Approaches using Thermal Tolerance Measures: A case study using *Drosophilids*

We integrated static lethal limits ($CT_{max}$) collected by @Heerwaarden2021 and XX with dynamic lethal limit measures from TDT curces [i.e., slope ($z^{TDT}$) and intercept ($CT_{max}^{TDT}$)] estimated by XX. We also collected data to estimate TDT curve parameters for eight additonal species following protocols outlined in @Orsted2022. Critical fertility limits for *Drosophilids* were compiled from @Parratt2021 and @Heerwaarden2021. These data were integrated with data on lethal limits at the species-level. When the same species were sampled multiple times we retained within species variation in models. We then fit the following model in *MCMCglmm* [@]:

$$
\begin{bmatrix}
 CT_{max} \\
 CT_{max}^{TDT} \\
 z^{TDT} \\
 CFL
\end{bmatrix} = u_{i,k} + p_{i,k} + sp_{i,k} + e_{i,k}
$$
